Privacy Aware Mobile Sensor Applications

### 2015-01-31 paper.tex
Added abstract draft.

\title{Privacy Analysis of a Mobile Sensor Application}
\author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
\institute{University of Koblenz-Landau}
\maketitle

@@

* The system has been built in Live+Gov EU project
* Analysis of Privacy aspects following a structured approach
* Derive eight privacy recommendations
* Some of them have been implemented in the service

@@

The rise of mobile smart-phones equipped with a wide range of sensors
and the abundance mobile internet connectivity has paved the way to a
whole new generation of mobile services which ease the daily life of
citizens of the western developed world. Apps like Google
Maps\footnote{\url{http://google.com/maps}} allows the user to
navigate effectively in unknown cities. Others track sports activities
in order to optimize training plans or engage in virtual competitions
(e.g. Strava\footnote{Citation Needed}).

Besides the immediate benefits, citizens using these services, reveal
a wealth of which can be used to draw a very detailed picture of the
private life of the citizen. For instance, GPS location tracking can
be used to infer shopping habits or associations with political groups
(when meetings are attended). Accelerometer data can be used to detect
medical conditions like walking disabilities. All this data is highly
sensitive to the citizens privacy and can be used against the citizen
if it falls into the wrong hands.

@@

Travel surveys are regularly conducted by providers of public
transportation infrastructure in order to asses the current usage of
their services and inform future planning of e.g. bus or tram lines.
In this events a large number of citizens (e.g. 5000) is asked to keep
track of their travel patterns for a constraint period of time
(e.g. one month). Traditionally the travel patterns have been recorded
manually in the form of travel diaries. Modern smart-phones are
equipped with a wide range of sensors (in particular GPS) that allow
the recording of travel-patterns in a fully automated way.

Within the Live+Gov EU project, we have established a first
prototypical implementation of such an automated travel-survey
service. In this section we give a high level description of a generic
travel-survey service. This description along with the introduced
terminology will serve as a basis for our privacy analysis.

There are three stakeholders involved with the sysem: The
\emph{citizen}, the \emph{service provider} and the \emph{travel
  agency}. The service provider is the operator of all IT systems and
sells the travel-survey service to the travel agency. Figure
\ref{figure:Live+Gov Vulnerabilities} contains a schematic
visualization of the relation between these stakeholders.

>>>>

\title{Privacy Aware Mobile Sensor Applications}
\author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
\institute{University of Koblenz-Landau}
\maketitle

@@

In this article we analyze the privacy aspects of a mobile sensor
application used for recording urban travel patterns as part of a
travel-survey service. This service has been developed and
field-tested within the Live+Gov EU Project. The privacy analysis
follows a structured approach established in
\cite{Grimm:ItSecRefModel}. Eight privacy recommendations are derived,
and have already led to corresponding enhancements of the
travel-survey service.

@@

The rise of mobile smart-phones equipped with a wide range of sensors
and the abundance mobile internet connectivity has paved the way to a
whole new generation of mobile services which ease the daily life of
citizens of the western developed world. Apps like Google
Maps\footnote{\url{http://google.com/maps}} allows the user to
navigate effectively in unknown cities. Others
(e.g. Strava\footnote{Citation Needed}) track sports activities in
order to optimize training plans or engage in virtual competitions .

Besides the immediate benefits citizens using these services reveal
a wealth of which can be used to draw a very detailed picture of the
private life of the citizen. For instance, GPS location tracking can
be used to infer shopping habits or associations with political groups
(when meetings are attended). Accelerometer data can be used to detect
medical conditions like walking disabilities. All this data is highly
sensitive to the citizens privacy and can be used against the citizen
if it falls into the wrong hands.

@@

Travel surveys are regularly conducted travel agencies which provide
public transportation infrastructure. With these surveys the current
usage of the public transportation system is assessed and the insights
are used for future planning of bus, tram and subway lines, etc..  For
each survey a large number of citizens (e.g. 5000) is asked to keep
track of their travel patterns for a constraint period of time
(e.g. one month). Traditionally the travel patterns have been recorded
manually in the form of travel diaries. Modern smart-phones are
equipped with a wide range of sensors (in particular GPS) that allow
the recording of travel-patterns in a fully automated way.

Within the Live+Gov EU project, we have established a first
prototypical implementation of such an automated travel-survey
service. In this section we give a high level description of the
generic service architecture. This description along with the
introduced terminology will serve as a basis for our privacy analysis.

There are three stakeholders involved in the travel-survey service:
The \emph{citizen}, the \emph{service provider} and the \emph{travel
  agency}. The service provider is the operator of all IT systems and
offers the travel-survey service to the travel agency. Figure
\ref{figure:Live+Gov Vulnerabilities} contains a schematic
visualization of the relation between these stakeholders.
